\chapter*{Conclusion}
\addcontentsline{toc}{chapter}{Conclusion}


During the course of this internship, I had the valuable opportunity to delve into the world of network automation and configuration management using NETCONF and YANG in a virtual network environment with GNS3. Working with Cisco CSR1000v routers and Python's ncclient library, I explored the possibilities of automating network configurations and harnessing the power of YANG data modeling.

Throughout the internship, I successfully enabled NETCONF and YANG support on the Cisco CSR1000v routers, establishing secure connections to manage configurations. I gained insights into essential network services like DHCP, NAT, and SSH, enabling efficient communication and secure access to the routers.

The experience of configuring networks with NETCONF using Python code was particularly enlightening. Leveraging the ncclient library, I interacted with the routers programmatically, retrieving and modifying configurations with ease. The understanding of base NETCONF operations empowered me to tailor configurations to meet specific requirements effectively.

Furthermore, I tackled the challenge of subnet allocation, a critical aspect in managing a large institution's network. By assessing the IP address requirements of each router and allocating appropriate subnets, I ensured efficient resource utilization and minimized IP address wastage.

In conclusion, this internship has been an enriching journey, providing hands-on exposure to cutting-edge network automation technologies. The knowledge gained in working with NETCONF and YANG, combined with practical skills in Python programming and GNS3 simulations, has equipped me with valuable tools for future endeavors in network engineering and automation.

I am grateful for the guidance and support from my mentors and colleagues, which played a significant role in the successful execution of this internship project. The experience gained here has instilled in me a deeper passion for network automation, and I look forward to applying these skills to contribute to the advancement of network management in real-world scenarios.