\chapter{Background}
Since network management became an essence for computer networks, the development of related technology has always been coupled with protocol standardizations, which are mainly OSI-based CMIP and TCP/IP-based SNMP. On one hand, OSI-SM has been the most powerful technology but is complicated and expensive and relies on OSI protocols that have gone out of fashion. On the other hand, SNMP is the very solution that has been used by most of the industry, but fell victim to its own simplicity: its data modeling capabilities are rudimentary and it does not support configuration management well due to its lack of transaction capabilities.

With the development of computer networks in multiple dimensions (number of devices, time scale for configuration, etc), configuring large networks becomes an increasingly difficult work. A set of configuration management requirements for IP-based networks are then identified, focusing on network–wide configurations, which provide a level of abstraction above device-local configurations. In this case, the function of configuration data translator must be seriously considered. Other requirements consist of distinguishing between configuration state and operational state, providing primitives to support concurrency in transaction-oriented way, the persistence of configuration changes, security considerations, and so on.

XML-based configuration management is now under hot research, especially using NETCONF. Main characteristics of the NETCONF protocol are briefly introduced as follows, with a detailed specification in Reference.

NETCONF defines a simple mechanism, through which a network device can be managed, configuration data information can be retrieved, and new configuration data can be uploaded and manipulated. The paradigm that it uses is named Remote Procedure Call (RPC). A key aspect of NETCONF is that, it allows the functionality of the management protocol to closely mirror the native functionality of the device. And applications can access both the syntactic content and the semantic content of the device's native user interface. In addition, NETCONF allows a client to discover the set of protocol extensions supported by a server. These so-called “capabilities” permit the client to adjust its behavior to take advantage of the features exposed by the device.

NETCONF Data Modeling Language (netmod) Working Group (WG) proposed by the IETF aims in supporting the ongoing development of IETF and vendor-defined data models for NETCONF, since NETCONF needs a standard content layer and its specifications do not include a modeling language or accompanying rules that can be used to model the management information to be configured using NETCONF. The main purpose of the netmod WG is to provide a unified data modeling language to standardize the NETCONF content, by defining a “human-friendly” language and emphasizing readability and ease of use. The very language seems to be able to serve as the normative description of NETCONF data models. Thus from this point of view, this WG plans to use YANG as its starting point for this language.
In the context of this internship project, the GNS3 network simulation tool is utilized to deploy simple network topologies. GNS3 allows for the creation of network environments that closely resemble real-world networks. By utilizing GNS3, interns can gain practical experience in network deployment and configuration, providing a suitable environment for testing and evaluating the operation of NETCONF and its interaction with network devices.

By combining the practical use of GNS3 and the implementation of NETCONF, this internship project aims to explore the deployment and management of network devices using software-defined networking principles.